Performance comparison of Dask and Apache Spark on HPC systems

### 2022-02-18 latex/paper.tex
[UPDATE] Address paper comments.

\section{Introduction}
The rise in data sharing coupled with improved data collection technologies leads neuroimaging into the Big Data era\cite{ALFAROALMAGRO:18, van2014human, ConpPortal}.
While standard neuroimaging workflow engines, such as Nipype\cite{Nipype:11}, are well rounded to process standard compute-intensive pipelines,
they lack Big Data strategies (i.e., in-memory computing, data locality, and lazy-evaluation) to improve the performance of increasingly prevalent data-intensive pipelines.
A previous study\cite{hayot2019performance} noted the importance of in-memory computing and data locality to improve the performance of data-intensive pipelines.
In this work, we study how performance benchmarks generalize across Big Data engines in an HPC environment.
In particular, we study in-depth the performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16} for their suitability to process neuroimaging pipelines.
Our objective is to assess whether Spark or Dask has substantial performance benefits in processing data-intensive HPC pipelines.

@@

While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructures.
This paper focuses on the performance of Big Data engines in HPCs environments.
We installed the Slurm scheduler and Lustre file system~\cite{braam2019lustre} on a dedicated cluster to mimic a typical HPC system.
Lustre is a popular file system for HPC environments offering parallel file accesses.

@@

The above studies compared the performance of the engine at a high level.
In comparison, our study provides a detailed analysis of the performance differences and similarities between Dask and Spark for data-intensive applications.

@@

\section{Background}
\subsection{Lustre}
The Lustre file system is a parallel file system well suited for HPC environments.
Lustre offers a POSIX-compliant interface with the file system.
Installed in some of the most powerful supercomputers, the design of Lustre allows
scaling to thousands of nodes, petabytes of storage, and hundreds of gigabits per second of I/O throughput.

@@

@@

\section{Methods}
\subsection{Infrastructure}
We used the ``slashbin" cluster at Concordia University.
The cluster has eight compute nodes, four storage nodes, one login node, and one Lustre metadata node.
Each compute node has two 16-core Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz,
\SI{250}{\gibi\byte} of RAM, \SI{126}{\gibi\byte} of tmpfs,
% 6 $\times$ SSDs of \SI{447}{\gibi\byte} each with the XFS file system (no RAID enabled),
CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.

We configured Lustre to have 1 MGS and 4 OSS.
Each OSS has 11 OSTs (disks) of \SI{8.8}{\tebi\byte}.
The OSTs are configured to have a maximum of \SI{1000}{\mebi\byte} of dirty data written into the client page cache.
Both the MDT and OSTs are configured to a maximum of 64 concurrent RPCs in flight.
A timeout of \SI{100}{\second} is set for Lustre, and OSTs checksum are disabled.
Lustre \textit{2.14.0-1.el8} is installed on both the Lustre clients and servers.

@@

\subsection{Applications}
In this section, we describe the applications used to benchmark Dask and Spark.
We use three simple synthetic applications to have a deep understanding of the 
underlying behavior of the applications with simple I/O patterns, namely
\textit{Increment}, \textit{Multi-Increment}, and \textit{Histogram}.
Also, two neuroimaging applications are used to study more realistic 
applications, namely \textit{Kmeans} and \textit{BIDS App Example}.

@@

\subsubsection{Increment}
We adapted the increment application used in~\cite{hayot2019performance}.
This synthetic application reads blocks of the BigBrain from Lustre and
simulates computation by sleeping for a specified period (Fig.~\ref{fig:graph-increment}). To simulate
intermediate results, we repeat the sleep process for a configurable amount
of time. We prevent data caching of the blocks by incrementing their voxels
value by one after each sleep operation. Finally, we write the resulting
NIfTI image back to Lustre. This application allows us to study the engines
when their inputs are processed independently. The map-only pattern
mimics the processing of multiple independent subjects in parallel.
% \begin{figure}[!hb]
% 	\centering
% 	\includegraphics[height=0.5\columnwidth,
% 	angle=0]{figures/increment.png}
% 	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
% 	\label{fig:graph-increment}
% \end{figure}

@@

\section{Discussion}
\subsection{I/O bottleneck}
We used Lustre as shared file system as it is a standard solution in HPC systems.
The performance of our Lustre installation degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
In a previous study~\cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
In this paper, Lustre shows some scaling when increasing the number of nodes, although not linear.
This suggests that increasing the parallelism level of the file system, including through potential configuration optimizations, might help alleviate the I/O bottleneck observed.

@@

\subsection{Memory management}
Our Kmeans experiment shows a significant difference between Dask and Spark.
The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
This is unexpected, given that Dask and Spark are configured similarly and use the same amount of resources.
We think those issues arise from a combination of
(1) the reserved memory for the JVM heap space,
(2) the serialization of Python to Java kept in memory,
(3) Spark only using procesess compared to Dask using threads, thus limiting shared memory consumption,
and (4) different policies to persist data to disk when running out of memory.

When reducing the number of processes from 64 to 32 for Spark (effectively doubling the memory of each process), the total execution time became similar between Spark and Dask.
However, with only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
We suppose that given more memory per thread/process, Dask and Spark would then perform similarly.

@@

\bibliographystyle{wileyNJD-AMA}
\bibliography{reference}
\end{document}

>>>>

\section{Introduction}
The rise in data sharing coupled with improved data collection technologies leads neuroimaging into the Big Data era\cite{ALFAROALMAGRO2018400, van2014human, ConpPortal}.
While standard neuroimaging workflow engines, such as Nipype\cite{Nipype:11}, are well rounded to process standard compute-intensive pipelines,
they lack Big Data strategies (i.e., in-memory computing, data locality, and lazy-evaluation) to improve the performance of increasingly prevalent data-intensive pipelines.
A previous study\cite{8752675} noted the importance of in-memory computing and data locality to improve the performance of data-intensive pipelines.
In this work, we study how performance benchmarks generalize across Big Data engines in an HPC environment.
In particular, we study in-depth the performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16} for their suitability to process neuroimaging pipelines.
Our objective is to assess whether Spark or Dask has substantial performance benefits in processing data-intensive HPC pipelines.

@@

While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructures.
This paper focuses on the performance of Big Data engines in HPCs environments.
We installed the Slurm scheduler and Lustre file system~\cite{Braam2019TheLS} on a dedicated cluster to mimic a typical HPC system.
Lustre is a reference file system for HPC environments, used in over 60 of the TOP100 supercomputers~\cite{OpenSFS-lustre}, offering parallel file accesses.

@@

The above studies compared the performance of the engine at a high level.
In comparison, our study provides a detailed analysis of the performance differences
and similarities, for HPC, between Dask and Spark for data-intensive applications.

@@

\section{Background}
\subsection{Lustre}
The Lustre file system is a parallel file system of reference in HPC environments.
Lustre offers a POSIX-compliant interface with the file system.
Installed in some of the most powerful supercomputers, the design of Lustre allows
scaling to thousands of nodes, petabytes of storage, and hundreds of gigabits per second of I/O throughput.

@@

While other filesystem such as HDFS or cloud storage could be considered, we chose Lustre to best represent an HPC environment.
HDFS makes use of permanent local storage on compute node, which is usually not feasable in HPC.
Morevoer, HPC node rarely have access to internet, limiting the use of cloud storage.

@@

\section{Methods}
\subsection{Infrastructure}
We used the ``slashbin" cluster at Concordia University.
The cluster has eight compute nodes, four storage nodes, one login node, and one Lustre metadata node.
Each compute node has two 16-core Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz,
\SI{250}{\gibi\byte} of RAM, \SI{126}{\gibi\byte} of tmpfs,
% 6 $\times$ SSDs of \SI{447}{\gibi\byte} each with the XFS file system (no RAID enabled),
CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
All of the compute and storage nodes are inter-connected by a \SI{10}{\giga\bit\per\second} Ethernet link through a switch.
The login node, hosting an NFS, is also connected by a \SI{10}{\giga\bit\per\second} Ethernet link to the compute nodes.

We configured Lustre to have 1 MGS and 4 OSS.
Each OSS has 11 OSTs (disks) of \SI{8.8}{\tebi\byte}.
The OSTs are configured to have a maximum of \SI{1000}{\mebi\byte} of dirty data written into the client page cache.
Both the MDT and OSTs are configured to a maximum of 64 concurrent RPCs in flight.
A timeout of \SI{100}{\second} is set for Lustre, and OSTs checksum are disabled.
Lustre \textit{2.14.0-1.el8} is installed on both the Lustre clients and servers.
%  TODO Inode config??

Our system uses similar hardware as Compute Canada Beluga, and replicates as
close as possible it's configuration.
This let us make a fair comparison of Spark and Dask in the condition of a realistic
HPC system.

@@

\subsection{Applications}
In this section, we describe the applications used to benchmark Dask and Spark.
We use three simple synthetic applications to have a deep understanding of the 
underlying behavior of the applications with simple I/O patterns, namely
\textit{Increment}, \textit{Multi-Increment}, and \textit{Histogram}.
Also, two neuroimaging applications are used to study more realistic 
applications, namely \textit{Kmeans} and \textit{BIDS App Example}\cite{gorgolewski2017bids}.

@@

The scripts used for our experiements and the results generated are available on
GitHub at: \href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines/}.

\subsubsection{Increment}
We adapted the increment application used in~\cite{8752675}.
This synthetic application reads blocks of the BigBrain from Lustre and
simulates computation by sleeping for a specified period (Fig.~\ref{fig:graph-increment}). To simulate
intermediate results, we repeat the sleep process for a configurable amount
of time. We prevent data caching of the blocks by incrementing their voxels
value by one after each sleep operation. Finally, we write the resulting
NIfTI image back to Lustre. This application allows us to study the engines
when their inputs are processed independently. The map-only pattern
mimics the processing of multiple independent subjects in parallel.
% \begin{figure}[!hb]
% 	\centering
% 	\includegraphics[height=0.5\columnwidth,
% 	angle=0]{figures/increment.png}
% 	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
% 	\label{fig:graph-increment}
% \end{figure}

@@

\section{Discussion}
\subsection{I/O bottleneck}
We used Lustre as shared file system as it is a defacto solution in HPC systems.
The performance of our Lustre installation degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
In a previous study~\cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
In this paper, Lustre shows some scaling when increasing the number of nodes, although not linear.
This suggests that increasing the parallelism level of the file system, including through potential configuration optimizations, might help alleviate the I/O bottleneck observed.

@@

\subsection{Memory management}
Our Kmeans experiment shows a significant difference between Dask and Spark.
The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
This is unexpected, given that Dask and Spark are configured similarly and are allocated the same amount of resources.
We think those issues arise from a combination of
(1) the reserved memory for the JVM heap space,
(2) the serialization of Python to Java kept in memory,
(3) Spark only using procesess compared to Dask using threads, thus limiting shared memory consumption,
and (4) different policies to persist data to disk when running out of memory.

When reducing the number of processes from 64 to 32 for Spark (effectively doubling the memory of each process), the total execution time became similar between Spark and Dask.
That is, with only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
We suppose that given more memory per thread/process, Dask and Spark would then perform similarly.
For our experiments, we allocated all available memory on the compute nodes.
Therefore, it was not possible to allocate more resources to test this hypothesis.

@@

\bibliographystyle{wileyNJD-AMA}
\bibliography{paper}
\end{document}
